%% Copyright (C) 2011, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\section{Profile.Function}

A \lstinline|Function| object contains information about calls of a
given function during a profiling operation.

\subsection{Prototypes}

\begin{refObjects}
\item[Object]
\end{refObjects}

\subsection{Construction}

\lstinline|Function| objects are not meant to be cloned as they are
created by \refSlot[System]{profile} internal machinery.

\begin{urbiscript}
function Float.fact()
{
  if (this <= 1)
    this
  else
    this * (this - 1).fact;
}|;
var profile = System.profile(function() { 6.fact });
[00009050] Profile(
[:]  Yields              :      0
[:]  Total time      (us):    337
[:]  Wall-clock time (us):    337
[:]  Function calls      :     23
[:]  Max depth           :      8
[:]
[:]  .-------------------------------------------------------------------.
[:]  |   function   |   %    | cumulative |  self  |  calls  |   self    |
[:]  |              |        |    (us)    |  (us)  |         | (us/call) |
[:]  |--------------+--------+------------+--------+---------+-----------|
[:]  |         fact |  63.20 |        213 |    213 |       6 |        35 |
[:]  |           <= |  10.98 |        250 |     37 |       6 |         6 |
[:]  |            - |   9.20 |        281 |     31 |       5 |         6 |
[:]  |            * |   8.61 |        310 |     29 |       5 |         5 |
[:]  |              |   4.75 |        326 |     16 |       1 |        16 |
[:]  |     profiled |   3.26 |        337 |     11 |       1 |        11 |
[:]  '--------------'--------'------------'--------'---------'-----------'
[:])
profile.calls[0];
[00123833] Function('fact', 6, 213, 35)
\end{urbiscript}

\subsection{Slots}

\begin{urbiscriptapi}
\item[calls]%
  The number of times this function was called during the profiling.
\begin{urbiassert}
// Example continued from Construction section.
profile.calls[0].calls == 6;
\end{urbiassert}

\item[name]%
  The name of the function called.
\begin{urbiassert}
// Example continued from Construction section.
profile.calls[0].name == "fact";
\end{urbiassert}

\item[selfTimePer]%
  Average CPU time spent in one function call. It is computed as the ratio
  of \refSlot{selfTime} divided by \refSlot{calls}.
\begin{urbiassert}
// Example continued from Construction section.
do (profile.calls[0])
{
  selfTimePer == selfTime / calls;
}
\end{urbiassert}

\item[selfTime]%
  Total CPU time spent in all calls of the function.
\begin{urbiassert}
// Example continued from Construction section.
profile.calls[0].selfTime.isA(Float);
\end{urbiassert}

\end{urbiscriptapi}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
