%% Copyright (C) 2010, 2011, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\begin{partDescription}{part:platforms}
  {
    %
    This part covers the specific features of \urbi for some of the
    platforms it was ported to.  Environments not described in this
    part are covered in separate, stand-alone, documentations.
    %
  }
\ifthen{\boolean{nao}}{
\item[sec:nao]
  Nao is a humanoid robot Nao from Aldebaran Robotics.
}

\ifthen{\boolean{bioloid}}{
\item[sec:bioloid]
  Using the Bioloid robot construction kit with \urbi.
}

\ifthen{\boolean{nxt}}{
\item[sec:nxt] LEGO Mindstorms NXT is a programmable robotics
  kit released by Lego in July 2006, replacing the first-generation
  LEGO Mindstorms kit. The kit consists of 519 Technic pieces, 3 servo
  motors, 4 sensors (ultrasonic, sound, touch, and light), 7
  connection cables, a USB interface cable, and the NXT Intelligent
  Brick.
}

\ifthen{\boolean{jazz}}{
\item[sec:jazz]%
  Gostai Open Jazz is an entirely programmable robot. It is shipped with the
  \usdk to develop with Jazz.  }

\ifthen{\boolean{p3dx}}{
\item[sec:p3dx] Pioneer 3-DX8 is an agile, versatile intelligent
  mobile robotic platform updated to carry loads more robustly and to
  traverse sills more surely with high-performance current management
  to provide power when it's needed.}

\ifthen{\boolean{rmp}}{
\item[sec:segway-rmp] The Segway Robotic Mobility Platform is a
  robotic platform based on the Segway Personal Transporter.  }

\ifthen{\boolean{spykee}}{
\item[sec:spykee] The Spykee is a WiFi-enabled robot built by Meccano
  (known as Erector in the United States). It is equipped with a
  camera, speaker, microphone, and moves using two tracks.
}

\ifthen{\boolean{webots}}{
\item[sec:webots] Using Cyberbotics' Webots simulation environment
  with \urbi.
}
\end{partDescription}


%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
