%% Copyright (C) 2010, 2011, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\chapter{Segway RMP}
\label{sec:segway-rmp}

The Segway \dfn{RMP} (Robotic Mobility Platform) is a robotic platform based
on the Segway Personal Transporter.  \urbi for Segway RMP can be downloaded
from \url{http://www.urbiforge.org/index.php/Robots/Segway}.

\sectionObject{RMP}
The UObject RMP provides the interface from \urbi.

\subsection{Instantiation}

\begin{urbiunchecked}
var rmp = RMP.new(\var{period});
\end{urbiunchecked}

\var{period} is the period, in seconds, at which orders will be sent and
coders will be read.

\subsection{Slots}

\paragraph{Raw actuator variables}

\begin{urbiscriptapi}
\item[forwardSpeed] forward speed in meters per second.  Can be read to
  (behaves as a sensor), or written to (actuator).


\item[yawSpeed] yaw speed in radian per second.  Can be read and written
  to.
\end{urbiscriptapi}

\paragraph{Raw sensors}
Current speed and coder values.
\begin{urbiscriptapi}
\item[forward]


\item[left]


\item[leftSpeed]


\item[right]


\item[rightSpeed]


\item[yaw]
\end{urbiscriptapi}

\paragraph{Calibration}

\urbi unit per coder unit coefficients.

\begin{urbiscriptapi}
\item[forwardCoeff]


\item[forwardSpeedCoeff]


\item[leftCoeff]


\item[leftSpeedCoeff]


\item[rightCoeff]


\item[rightSpeedCoeff]


\item[yawCoeff]


\item[yawSpeedCoeff]
\end{urbiscriptapi}

The file \file{usegway.u} contains approximate values for meter, seconds,
radian units.


\section{GSRAPI compliance}

The file \file{rmp.u} will initialize the RMP UObject and give it a Gostai
Standard Robotic API compliant interface.


%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
