%% Copyright (C) 2009-2012, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\chapter{Functional Programming}
\label{sec:tut:functional}

\us support functional programming through first class functions and
lambda expressions.

\section{First class functions}

\us has first class functions, i.e., functions are regular values,
just like integers or strings. They can be stored in variables,  passed
as arguments to other functions, and so forth. For instance, you don't need
to write
\lstinline|function object.f(){/* ... */}| to insert a function in an
object, you can simply use \lstinline{setSlot}.

% FIXME: doesn't work, toplevel, puke, etc
\begin{urbiscript}[firstnumber=1]
var o = Object.clone|;
// Here we can use f as any regular value.
o.setSlot("m1", function () { echo("Hello") })|;
// This is strictly equivalent.
var o.m2 = function () { echo("Hello") }|;
o.m1;
[00000000] *** Hello
o.m2;
[00000000] *** Hello
\end{urbiscript}

This enables to write powerful pieces of code, like functions that
take function as argument. For instance, consider the \lstinline{all}
function: given a list and a function, it applies the function to each
element of the list, and returns whether all calls returned true. This
enables to check very simply if all elements in a list verify a
predicate.

\begin{urbiscript}
function all(list, predicate)
{
  for (var elt : list)
    if (!predicate(elt))
      return false;
  return true;
}|;
// Check if all elements in a list are positive.
function positive(x) { x >= 0 }|;
all([1, 2, 3], getSlot("positive"));
[00000000] true
all([1, 2, -3], getSlot("positive"));
[00000000] false
\end{urbiscript}

It turns out that \lstinline|all| already exists: instead of
\lstinline|all(\var{list}, \var{predicate})|, use
\lstinline|\var{list}.all(\var{predicate})|, see
\refSlot[RangeIterable]{all}.

\section{Lambda functions}

Another nice feature is the ability to write lambda functions, which
are anonymous functions. You can create a functional value as an
expression, without naming it, with the syntax shown below.

\begin{urbiscript}
// Create an anonymous function
function (x) {x + 1}|;
// This enable to easily pass function
// to our "all" function:
[1, 2, 3].all(function (x) { x > 0});
[00000000] true
\end{urbiscript}

In fact, the \lstinline{function} construct we saw earlier is only a
shorthand for a variable assignment.

\begin{urbiunchecked}
// This ...
function obj.f (/*...*/) {/*...*/};
// ... is actually a shorthand for:
const var obj.f = function (/*...*/) {/* ... */};
\end{urbiunchecked}

% This should maybe be outside the functional section. It is also
% incomplete.
\section{Lazy arguments}

Most popular programming languages use strict arguments evaluation:
arguments are evaluated before functions are called. Other languages use
lazy evaluation: argument are evaluated by the function only when needed. In
\us, evaluation is strict by default, but you can ask a function not to
evaluate its arguments, and do it by hand. This works by not specifying
formal arguments. The function is provided with a \lstinline{call} object
that enables you to evaluate arguments.

\begin{urbiscript}
// Note the lack of formal arguments specification
function first
{
  // Evaluate only the first argument.
  call.evalArgAt(0);
}|;
first(echo("first"), echo("second"));
[00000000] *** first
function reverse
{
  call.evalArgAt(1);
  call.evalArgAt(0);
}|;
reverse(echo("first"), echo("second"));
[00000000] *** second
[00000000] *** first
\end{urbiscript}

A good example are logic operators. Although \Cxx is a strict language, it
uses a few logic operators. For instance, the logical and (\lstinline{&&})
does not evaluate its right operand if the left operand is false (the result
will be false anyway).

\us logic operator mimic this behavior. The listing below shows how one can
implement such a behavior.

\begin{urbiscript}
function myAnd
{
  if (call.evalArgAt(0))
    call.evalArgAt(1)
  else
    false;
}|;

function f()
{
  echo("f executed");
  return true;
}|;

myAnd(false, f());
[00000000] false

myAnd(true, f());
[00000000] *** f executed
[00000000] true
\end{urbiscript}

%\section{Pattern matching}

%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
