%% Copyright (C) 2009-2010, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\section{System.Platform}
A description of the platform (the computer) the server is running on.

\subsection{Prototypes}
\begin{itemize}
\item \refObject{Object}
\end{itemize}

\subsection{Slots}
\begin{urbiscriptapi}
\item[host] The type of system \usdk runs on.  Composed of the CPU, the
  vendor, and the OS.
\begin{urbiassert}
System.Platform.host ==
  "%s-%s-%s" % [System.Platform.hostCpu,
                System.Platform.hostVendor,
                System.Platform.hostOs];
\end{urbiassert}

\item[hostAlias] The name of the system \usdk runs on as the person who
  compiled it decided to name it.  Typically empty, it is fragile to depend
  on it.
\begin{urbiassert}
System.Platform.hostAlias.isA(String);
\end{urbiassert}

\item[hostCpu] The CPU type of system \usdk runs on.  The following values
  are those for which Gostai provides binary builds.
\begin{urbiassert}
System.Platform.hostCpu in ["i386", "i686", "x86_64"];
\end{urbiassert}

\item[hostOs] The OS type of system \usdk runs on.  For instance
  \lstinline|darwin9.8.0| or \lstinline|linux-gnu| or \lstinline|mingw32|.

\item[hostVendor] The vendor type of system \usdk runs on.  The following
  values are those for which Gostai provides binary builds.
\begin{urbiassert}
System.Platform.hostVendor in ["apple", "pc", "unknown"];
\end{urbiassert}

\item[isWindows] Whether running under Windows.
\begin{urbiassert}
System.Platform.isWindows in [true, false];
\end{urbiassert}

\item[kind] Either \code{"POSIX"} or \code{"WIN32"}.
\begin{urbiassert}
System.Platform.kind in ["POSIX", "WIN32"];
\end{urbiassert}
\end{urbiscriptapi}

%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../../urbi.dict"
%%% fill-column: 76
%%% End:
