\section{Boolean}

There is no object \lstinline|Boolean| in \us, but two specific
objects \lstindex{true} and \lstindex{false}.  They are the result of
all comparison statement.


\subsection{Prototypes}

The objects \lstinline|true| and \lstinline|false| have the following
prototype.

\begin{itemize}
\item \refObject{Singleton}
\end{itemize}

\subsection{Construction}

There are no constructors, use \lstinline|true| and \lstinline|false|.
Since they are singletons, \lstinline|clone| will return themselves,
not new copies.

\begin{urbiscript}
assert(true);
assert(!false);
assert(2 < 6 === true);
assert(true.new === true);
assert(6 < 2 === false);
\end{urbiscript}

\subsection{Truth Values}
\label{sec:truth}

As in many programming languages, conditions may be more than only
\lstinline|true| and \lstinline|false|.  Whether some value is
considered as true depends on the type of \lstinline|this|.  Actually,
by default objects as considered ``true'', objects evaluating to
``false'' are the exception:
\begin{itemize}
\item \lstinline|false|, \lstinline|nil|\\
  \lstinline|false|.
\item \lstinline|void|\\
  raise an error.
\item lstinline|Float|\\
  \lstinline|false| iff null (\refObject{Float}).
\item \lstinline|Dictionary|, \lstinline|List|,  \lstinline|String|\\
  \lstinline|false| iff empty (\refObject{Dictionary},
  \refObject{List}, \refObject{String}).
\item otherwise\\
  \lstinline|true|.
\end{itemize}

The method \lstinline|Object.asBool| is in charge of converting some
arbitrary value into a Boolean.
\begin{urbiscript}[firstnumber=last]
assert(Global.asBool == true);
assert(nil.asBool ==    false);
void.asBool;
[00000421:error] !!! unexpected void
\end{urbiscript}


\subsection{Slots}

\begin{itemize}
\item \lstinline|'&&'(\var{that})|\\
  Short-circuiting logical and. If \lstinline|this| is
  \lstinline|true| evaluate and return \var{that}.  If
  \lstinline|this| is \lstinline|false|, return itself without
  evaluating \var{that}.
\begin{urbiscript}[firstnumber=last]
assert((true && 2) == 2);
assert((false && 1 / 0) == false);
\end{urbiscript}

\item \lstinline+'||'(\var{that})+\\
  Short-circuiting logical or. If \lstinline|this| is
  \lstinline|false| evaluate and return \var{that}.  If
  \lstinline|this| is \lstinline|true|, return itself without
  evaluating \var{that}.
\begin{urbiscript}[firstnumber=last]
assert((true || 1/0) == true);
assert((false || 2)  == 2);
\end{urbiscript}

\item \lstinline|'!'|\\
  Logical negation. If \lstinline|this| is \lstinline|false| return
  \lstinline|true| and vice-versa.
\begin{urbiscript}[firstnumber=last]
assert(!true == false);
assert(!false == true);
\end{urbiscript}

\item \lstinline|asBool|\\
  Identity.
\begin{urbiscript}[firstnumber=last]
assert(true.asBool ==  true);
assert(false.asBool == false);
\end{urbiscript}
\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% End:
