%% Copyright (C) 2009-2010, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\section{Channel}
Returning data, typically asynchronous, with a label so that the
``caller'' can find it in the flow.

\subsection{Prototypes}

\begin{itemize}
\item \refObject{Object}
\end{itemize}

\subsection{Construction}

Channels are created like any other object. The constructor must be
called with a string which will be the label.

\begin{urbiscript}[firstnumber=1]
var ch1 = Channel.new("my_label");
[00000201] Channel_0x7985810

ch1 << 1;
[00000201:my_label] 1

var ch2 = ch1;
[00000201] Channel_0x7985810

ch2 << 1/2;
[00000201:my_label] 0.5
\end{urbiscript}

\subsection{Slots}

\begin{urbiscriptapi}
\item['<<'](<value>)%
  Send \var{value} to \this tagged by its label if non-empty.

\begin{urbiscript}
Channel.new("label") << 42;
[00000000:label] 42

Channel.new("") << 51;
[00000000] 51
\end{urbiscript}

\item[echo](<value>)%
  Same as \lstinline|lobby.echo(\var{value}, name)|, see
  \refSlot[Lobby]{echo}.

\begin{urbiscript}
Channel.new("label").echo(42);
[00000000:label] *** 42

Channel.new("").echo("Foo");
[00000000] *** Foo
\end{urbiscript}

\item[enabled] Whether the Channel is enabled.  Disabled Channels
  produce no output.
\begin{urbiscript}
var c = Channel.new("")|;

c << "enabled";
[00000000] "enabled"

c.enabled = false|;
c << "disabled";

c.enabled = true|;
c << "enabled";
[00000000] "enabled"
\end{urbiscript}

\item[Filter] Filtering channel.

The Filter channel outputs text that can be parsed without error by the liburbi.
It does this by filtering types not handled by the liburbi, and displaying
them using \refSlot{echo}.

\begin{urbiscript}
// Use a filtering channel on our lobby output.
topLevel = Channel.Filter.new("")|;
// liburbi knows about List, Dictionary, String and Float, so standard display.
[1, "foo", ["test" => 5]];
[00000001] [1, "foo", ["test" => 5]]
// liburbi does not know 'lobby', so it is escaped with echo:
lobby;
[00000002] *** Lobby_0xADDR
// The following list contains a function which is not handled by liburbi, so
// it gets escaped too.
[1, function () {}];
[00000003] *** [1, function () {}]
// Restore default display to see the difference.
topLevel = Channel.topLevel|;
// The echo is now gone.
[1, function () {}];
[00001758] [1, function () {}]
\end{urbiscript}

\item[quote] Whether the strings are output escaped (the default)
  instead of raw strings.
\begin{urbiscript}
var d = Channel.new("")|;

assert(d.enabled);
d << "A \"String\"";
[00000000] "A \"String\""

d.quote = false|;
d << "A \"String\"";
[00000000] A "String"
\end{urbiscript}

\item[name] The name of the Channel, used to label the output.
\begin{urbiscript}
assert
{
  Channel.new("").name == "";
  Channel.new("foo").name == "foo";
};
\end{urbiscript}

\item[null] A predefined stream whose \lstinline|enabled| is
  \lstinline|false|.
\begin{urbiscript}
Channel.null << "Message";
\end{urbiscript}


\item[topLevel] A predefined stream for regular output.  Strings are
  output escaped.
\begin{urbiscript}
Channel.topLevel << "Message";
[00015895] "Message"
Channel.topLevel << "\"quote\"";
[00015895] "\"quote\""
\end{urbiscript}

\item[warning] A predefined stream for warning messages.  Strings sent
  to it are not escaped.
\begin{urbiscript}
Channel.warning << "Message";
[00015895:warning] Message
Channel.warning << "\"quote\"";
[00015895:warning] "quote"
\end{urbiscript}
\end{urbiscriptapi}

%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
