%% Copyright (C) 2009-2010, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\section{Enumerated}

Prototype of enumerated types.

\subsection{Prototypes}
\begin{refObjects}
\item[Object]
\end{refObjects}

\subsection{Examples}

See \autoref{sec:lang:enum}.

\subsection{Construction}
\label{sec:stdlib:event:ctor}

An \lstinline{Enumerated} is created with two arguments: the name of the
enumerated type, and the list of possible values. Most of the time, it is a
good idea to store it in a variable with the same name.

\begin{urbiscript}[firstnumber=1]
var Direction = Enumerated.new("Direction", ["up", "down", "left", "right"]);
[00000001] Direction
Direction.up;
[00000002] up
\end{urbiscript}

You might however be more familiar with the following syntax, which is an
equivalent syntactic sugar.

\begin{urbiunchecked}
enum Direction
{
  up,
  down,
  left,
  right
};
[00000001] Direction
\end{urbiunchecked}

\subsection{Slots}
\begin{urbiscriptapi}

\item[asList]
  Synonym for \refSlot[Enumarated]{values}.

\item[each]
  Iterate through the possible values. Enables the use of the
  \lstinline{for} iteration.
  \begin{urbiscript}
for (var d in Direction)
  echo(d);
[00000001] *** up
[00000001] *** down
[00000001] *** left
[00000001] *** right
  \end{urbiscript}

\item[name]
  The name of the enum, as given at construction time.
  \begin{urbiscript}
Direction.name;
[00000001] "Direction"
  \end{urbiscript}

\item[size]
  Number of possible values in the enumarated type. Equivalent to
  \lstinline{values.size}.
  \begin{urbiscript}
Direction.size;
[00000001] 4
  \end{urbiscript}

\item[values]
  The list of values.
  \begin{urbiscript}
Direction.values;
[00000001] [up, down, left, right]
  \end{urbiscript}

\end{urbiscriptapi}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
