%% Copyright (C) 2009-2011, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\section{Lazy}

\dfn{Lazies} are objects that hold a lazy value, that is, a not yet
evaluated value. They provide facilities to evaluate their content only once
(\dfn{memoization}) or several times. Lazy are essentially used in call
messages, to represent lazy arguments, as described in
\autorefObject{CallMessage}.

\subsection{Examples}

\subsubsection{Evaluating once}

One usage of lazy values is to avoid evaluating an expression unless it's
actually needed, because it's expensive or has undesired side effects. The
listing below presents a situation where an expensive-to-compute value
(\lstinline|heavy_computation|) might be needed zero, one or two times. The
objective is to save time by:

\begin{itemize}
\item Not evaluating it if it's not needed.
\item Evaluating it only once if it's needed once or twice.
\end{itemize}

We thus make the wanted expression lazy, and use the \lstinline|value|
method to fetch its value when needed.

\begin{urbiscript}[firstnumber=1]
// This function supposedly performs expensive computations.
function heavy_computation()
{
  echo("Heavy computation");
  return 1 + 1;
}|;

// We want to do the heavy computations only if needed,
// and make it a lazy value to be able to evaluate it "on demand".
var v = Lazy.new(closure () { heavy_computation() });
[00000000] heavy_computation()
/* some code */;
// So far, the value was not needed, and heavy_computation
// was not evaluated.
/* some code */;
// If the value is needed, heavy_computation is evaluated.
v.value;
[00000000] *** Heavy computation
[00000000] 2
// If the value is needed a second time, heavy_computation
// is not reevaluated.
v.value;
[00000000] 2
\end{urbiscript}

\subsubsection{Evaluating several times}

Evaluating a lazy several times only makes sense with lazy arguments and
call messages. See example with call messages in
\autoref{sec:std-callmsg-examples-several}.


\subsection{Caching}

\refObject{Lazy} is meant for functions without argument.  If you need
\dfn{caching} for functions that depend on arguments, it is straightforward
to implement using a \refObject{Dictionary}.  In the future \us might
support dictionaries whose indices are not only strings, but in the
meanwhile, convert the arguments into strings, as the following sample
object demonstrates.

\begin{urbiscript}
class UnaryLazy
{
  function init(f)
  {
    results = [ => ];
    func = f;
  };
  function value(p)
  {
    var sp = p.asString;
    if (results.has(sp))
      return results[sp];
    var res = func(p);
    results[sp] = res |
    res
  };
  var results;
  var func;
} |
// The function to cache.
var inc = function(x) { echo("incing " + x) | x+1 } |
// The function with cache.
// Use "getSlot" to get the function, not its evaluation.
var p = UnaryLazy.new(getSlot("inc"));
[00062847] UnaryLazy_0x78b750
p.value(1);
[00066758] *** incing 1
[00066759] 2
p.value(1);
[00069058] 2
p.value(2);
[00071558] *** incing 2
[00071559] 3
p.value(2);
[00072762] 3
p.value(1);
[00074562] 2
\end{urbiscript}

\subsection{Prototypes}

\begin{refObjects}
\item[Comparable]
\end{refObjects}

\subsection{Construction}

Lazies are seldom instantiated manually. They are mainly created
automatically when a lazy function call is made (see
\autoref{sec:us-fun-callmsg}). One can however create a lazy value with the
standard \lstinline|new| method of \lstinline|Lazy|, giving it an
argument-less function which evaluates to the value made lazy.

\begin{urbiscript}
Lazy.new(closure () { /* Value to make lazy */ 0 });
[00000000] 0
\end{urbiscript}

\subsection{Slots}

\begin{urbiscriptapi}
\item[==](<that>)%
  Whether \this and \var{that} are the same source code and value (an not
  yet evaluated Lazy is never equal to an evaluated one).
\begin{urbiassert}
Lazy.new(closure () { 1 + 1 }) == Lazy.new(closure () { 1 + 1 });
Lazy.new(closure () { 1 + 2 }) != Lazy.new(closure () { 2 + 1 });
\end{urbiassert}
\begin{urbiscript}
{
  var l1 = Lazy.new(closure () { 1 + 1 });
  var l2 = Lazy.new(closure () { 1 + 1 });
  assert (l1 == l2);
  l1.eval;
  assert (l1 != l2);
  l2.eval;
  assert (l1 == l2);
};
\end{urbiscript}

\item[asString]
  The conversion to \refObject{String} of the body of a non-evaluated
  argument.
\begin{urbiassert}
Lazy.new(closure () { echo(1); 1 }).asString == "echo(1);\n1";
\end{urbiassert}

\item[eval]%
  Force the evaluation of the held lazy value. Two calls to \refSlot{eval}
  will systematically evaluate the expression twice, which can be useful to
  duplicate its side effects.

\item[value]%
  Return the held value, potentially evaluating it before. \refSlot{value}
  performs memoization, that is, only the first call will actually evaluate
  the expression, subsequent calls will return the cached value. Unless you
  want to explicitly trigger side effects from the expression by evaluating
  it several time, this should be preferred over \lstinline|eval| to avoid
  evaluating the expression several times uselessly.
\end{urbiscriptapi}


%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
