%% Copyright (C) 2011, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\section{Profile}

A \lstinline|Profile| object contains information about the efficiency of a
piece of code.

\subsection{Example}

\subsubsection{Basic profiling}

One can profile a piece of code with the \refSlot[System]{profile} function.

\begin{urbiscript}
var profile = System.profile(function() { echo("foo") });
[00000001] *** foo
[00001672] Profile(
[:]  Yields:                    0
[:]  Total time (us):         112
[:]  Wall-clock time (us):    112
[:]  Function calls:           14
[:]  Max depth:                 5
[:]
[:]  .-------------------------------------------------------------------.
[:]  |   function   |   %    | cumulative |  self  |  calls  |   self    |
[:]  |              |        |    (us)    |  (us)  |         | (us/call) |
[:]  |--------------+--------+------------+--------+---------+-----------|
[:]  |         send |  24.11 |         27 |     27 |       1 |        27 |
[:]  |         echo |  14.29 |         43 |     16 |       1 |        16 |
[:]  |        apply |  10.71 |         55 |     12 |       1 |        12 |
[:]  |        apply |   9.82 |         66 |     11 |       1 |        11 |
[:]  |            + |   8.93 |         76 |     10 |       2 |         5 |
[:]  |   <profiled> |   6.25 |         83 |      7 |       1 |         7 |
[:]  |     asString |   6.25 |         90 |      7 |       1 |         7 |
[:]  |        lobby |   5.36 |         96 |      6 |       2 |         3 |
[:]  | getSlotValue |   3.57 |        104 |      4 |       1 |         4 |
[:]  |            + |   2.68 |        107 |      3 |       1 |         3 |
[:]  |        Lobby |   1.79 |        112 |      2 |       2 |         1 |
[:]  '--------------'--------'------------'--------'---------'-----------'
[:])
\end{urbiscript}

The result is a \refObject{Profile} object that contains information about
which functions where used when evaluating the given code, how many time they
were called, how much time was spent in them, \ldots Lines are sorted by
decreasing ``self time''.  Note that the \lstinline|<profiled>| special
function stands for the function given in parameter.  Every line is
represented by a \refObject{Profile.Function} object, see its documentation
for the meaning of every column.

\subsubsection{Asynchronous profiling}

If the profiled code spawns asynchronous tasks via \lstinline|detach| or
\lstinline|at| for instance, additional statistics will be included in the
resulting \refObject{Profile} every time the detached code is executed. This
is extremely useful to profile asynchronous code based on \lstinline|at| for
instance.

\begin{urbiscript}
var x = false|;
// Make sure 'x' is visible whoever the caller of 'profiled' is.
import this.*;
function profiled()
{
  at (x)
    echo("true")
  onleave
    echo("false")
}|;

// This is the profiling for the creation of the 'at'. Note that the
// condition was already evaluated once, to see whether it should trigger
// immediately.
var profile_async = System.profile(getSlotValue("profiled"));
[00000000] Profile(
[:]  Yields:                    0
[:]  Total time (us):          73
[:]  Wall-clock time (us):     73
[:]  Function calls:           11
[:]  Max depth:                 4
[:]
[:]  .-------------------------------------------------------------------.
[:]  |   function   |   %    | cumulative |  self  |  calls  |   self    |
[:]  |              |        |    (us)    |  (us)  |         | (us/call) |
[:]  |--------------+--------+------------+--------+---------+-----------|
[:]  |   <profiled> |  46.58 |         34 |     34 |       1 |        34 |
[:]  |    at: { x } |  21.92 |         50 |     16 |       1 |        16 |
[:]  |      onEvent |  15.07 |         61 |     11 |       1 |        11 |
[:]  |        clone |   9.59 |         68 |      7 |       2 |     3.500 |
[:]  |            x |   2.74 |         70 |      2 |       2 |         1 |
[:]  |          new |   2.74 |         72 |      2 |       2 |         1 |
[:]  |         init |   1.37 |         73 |      1 |       2 |     0.500 |
[:]  '--------------'--------'------------'--------'---------'-----------'
[:])

// Trigger the at twice.
x = true|;
[00106213] *** true
x = false|;
[00172119] *** false

// The profile now includes additional statistic about the evaluations of
// the condition and the bodies of the at.
profile_async;
[00178623] Profile(
[:]  Yields:                    2
[:]  Total time (us):         251
[:]  Wall-clock time (us):    251
[:]  Function calls:           33
[:]  Max depth:                 4
[:]
[:]  .-------------------------------------------------------------------.
[:]  |   function   |   %    | cumulative |  self  |  calls  |   self    |
[:]  |              |        |    (us)    |  (us)  |         | (us/call) |
[:]  |--------------+--------+------------+--------+---------+-----------|
[:]  |   <profiled> |  13.55 |         34 |     34 |       1 |        34 |
[:]  |        event |  11.55 |         63 |     29 |       1 |        29 |
[:]  |         send |  11.16 |         91 |     28 |       2 |        14 |
[:]  |        event |  10.76 |        118 |     27 |       1 |        27 |
[:]  |    at: { x } |  10.76 |        145 |     27 |       3 |         9 |
[:]  |        clone |   5.98 |        183 |     15 |       4 |     3.750 |
[:]  |         echo |   5.58 |        212 |     14 |       2 |         7 |
[:]  |      onEvent |   4.38 |        223 |     11 |       1 |        11 |
[:]  |            + |   1.99 |        228 |      5 |       4 |     1.250 |
[:]  |          new |   1.99 |        233 |      5 |       4 |     1.250 |
[:]  |     asString |   0.80 |        242 |      2 |       2 |         1 |
[:]  |         init |   0.80 |        248 |      2 |       4 |     0.500 |
[:]  |            x |      0 |        251 |      0 |       1 |         0 |
[:]  |            x |      0 |        212 |      0 |       3 |         0 |
[:]  '--------------'--------'------------'--------'---------'-----------'
[:])
\end{urbiscript}

%% See issue #4969.
Note that part of the internal machinery shows in these figures (and left
visible on purpose).  For instance the three additional calls to
\lstinline{new} correspond to the creation of the \lstinline{changed} event.

\subsection{Prototypes}

\begin{refObjects}
\item[Object]
\end{refObjects}

\subsection{Construction}

\lstinline|Profile| objects are not meant to be cloned as they are created
by \refSlot[System]{profile} internal machinery.

\subsection{Slots}

\begin{urbiscriptapi}
\item[calls]%
  Return a \refObject{List} of \refObject{Profile.Function} objects. Each
  element of this list describes, for a given function, statistics about how
  many times it is called and how much time is spent in it.


\item[Function]
  See \refObject{Profile.Function}.
\begin{urbiscript}
Profile.Function;
[00004990] Function('', 0, 0, nan)
\end{urbiscript}


\item[maxFunctionCallDepth]%
  The maximum function call depth reached.
\begin{urbiassert}
// Example continued from Construction.
profile.maxFunctionCallDepth == 5;
\end{urbiassert}


\item[totalCalls]%
  The total number of function calls made.
\begin{urbiassert}
// Example continued from Construction.
profile.totalCalls == 14;
\end{urbiassert}


\item[totalTime]%
  The total CPU time. It can be higher than the wall clock time on
  multi-core processors for instance.


\item[wallClockTime]%
  The time spent between the beginning and the end as if measured on a wall
  clock.


\item[yields]%
  The scheduler has to execute many coroutines in parallel. A coroutine
  yields when it gives the opportunity to an other to be executed until this
  one yields and so on\dots This slot contains the number of scheduler
  yields.
\begin{urbiassert}
// Example continued from Construction.
profile.yields == 0;
\end{urbiassert}
\end{urbiscriptapi}


%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
