%% Copyright (C) 2009-2011, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\section{RangeIterable}

This object is meant to be used as a prototype for objects that support an
\lstinline|asList| method, to use range-based \lstinline|for| loops
(\autoref{sec:lang:for:each}).

\subsection{Prototypes}

\begin{refObjects}
\item[Object]
\end{refObjects}

\subsection{Slots}

\begin{urbiscriptapi}
\item[all](<fun>)
  % FIXME: link to predicate glossary entry
  Return whether all the members of the target verify the predicate
  \var{fun}.

\begin{urbiassert}
// Are all elements positive?
! [-2, 0, 2, 4].all(function (e) { e > 0 });
// Are all elements even?
[-2, 0, 2, 4].all(function (e) { e % 2 == 0 });
\end{urbiassert}

\item[any](<fun>)
  % FIXME: link to predicate glossary entry
  Whether at least one of the members of the target verifies the
  predicate \var{fun}.

\begin{urbiassert}
// Is there any even element?
! [-3, 1, -1].any(function (e) { e % 2 == 0 });
// Is there any positive element?
[-3, 1, -1].any(function (e) { e > 0 });
\end{urbiassert}

\item[each](<fun>)%
  Apply the given functional value \var{fun} on all ``members'',
  sequentially.  Corresponds to range-\lstinline|for| loops.
\begin{urbiscript}[firstnumber=1]
class range : RangeIterable
{
  var asList = [10, 20, 30];
}|;
for (var i : range)
  echo (i);
[00000000] *** 10
[00000000] *** 20
[00000000] *** 30
\end{urbiscript}

\item['each&'](<fun>)%
  Apply the given functional value \var{fun} on all ``members'', in
  parallel, starting all the computations simultaneously.  Corresponds
  to range-\lstinline|for&| loops.
\begin{urbiscript}
{
  var res = [];
  for& (var i : range)
    res << i;
  assert(res.sort == [10, 20, 30]);
};
\end{urbiscript}

\item \lstinline+'each|'(\var{fun})+%  Better done in master than 2.7.
  Apply the given functional value \var{fun} on all ``members'', with
  tight sequentially.  Corresponds to range-\lstinline'for|' loops.
\begin{urbiscript}
{
  var res = [];
  for| (var i : range)
    res << i;
  assert(res == [10, 20, 30]);
};
\end{urbiscript}
\end{urbiscriptapi}

%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
