\chapter*{Introduction}

\section*{Status of this document}

This document defines the specifications of the \us language version
2.0. It defines the expected behavior from the \us interpreter, the
standard library, and the \sdk. It can be used to check whether some code
is valid, or browse \us or \Cxx \api for a desired feature. Random reading
can also provide you with
advanced knowledge or subtleties about some \us aspects.

This document is not an \us tutorial; it is not structured in a
progressive manner and is too
detailed. Think of this document as a dictionary: one does not
learn a foreign language by reading a dictionary. The \us Tutorial, or
the live \us tutorial built in the interpreter are good introductions
to \us.

This document does not aim at giving advanced programming
techniques. Its only goal is to define the language and its
libraries.

\section*{\us}

\dfn[urbiscript@\us]{\us} is an interpreted language designed for highly
interactive and parallel behavior programming. %FILLME

\section*{Structure of this document}

This document structured as follows:

\begin{description}
  \newcommand{\xitem}[2]{\item[\autoref{#1} --- #2]~\\}
\xitem{sec:tools}{Tools specifications}%
  Presentation and usage of the different tools available with the
  \urbi framework related to \us, such as the \urbi server, the
  command line client, \umake, \ldots

\xitem{sec:lang}{\us language specifications}%
  Core constructs of the language and their behavior.

\xitem{sec:stdlib}{\us standard library specifications}%
  Listing of all classes and methods provided in the standard library.

\xitem{sec:sdk}{\urbi \sdk specifications}%
  The \urbi software development kit that enable to
  interact with \urbi from \Cxx.
\end{description}

\input{specs/tools}
\FloatBarrier
\input{specs/lang}
\FloatBarrier
\input{specs/stdlib}
\FloatBarrier

\chapter{\urbi SDK specifications}
\label{sec:sdk}
