%% Copyright (C) 2010, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\section{StackFrame}

This class is meant to record backtrace (see \refSlot[Exception]{backtrace})
information.

For convenience, all snippets of code are supposed to be run after these
function definitions.  In this code, the \lstinline|getStackFrame| function
is used to get the first \lstinline|StackFrame| of an exception backtrace.
Backtrace of \refObject{Exception} are filled with \lstinline|StackFrame|s
when the is thrown.

\begin{urbiscript}
//#push 1 "foo.u"
function inner () { throw Exception.new("test") }|;

function getStackFrame()
{
  try
  {
    inner
  }
  catch(var e)
  {
    e.backtrace[0]
  };
}|;
//pop
\end{urbiscript}

\experimental{}

%% \subsection{Prototypes}
%% \begin{refObjects}
%% \item[Orderable]
%% \item[Comparable]
%% \end{refObjects}

\subsection{Construction}

\lstinline|StackFrame| are not made to be manually constructed.  The
initialization function expect 2 arguments, which are the name of the called
function and the \refObject{Location} from which it has been called.

\begin{urbiscript}
StackFrame.new("inner",
  Location.new(
    Position.new("foo.u", 7, 5),
    Position.new("foo.u", 7, 10)
  )
);
[00000001] foo.u:7.5-9: inner
\end{urbiscript}

\subsection{Slots}

\begin{urbiscriptapi}
\item[name] \refObject{String}, representing the name of the called
  function.
\begin{urbiscript}
getStackFrame.name;
[00000002] "inner"
\end{urbiscript}

\item[location] \refObject{Location} of the function call.
\begin{urbiscript}
getStackFrame.location;
[00000003] foo.u:7.5-9
\end{urbiscript}

\item[asString] Clean display of the call location.
\begin{urbiscript}
getStackFrame;
[00000004] foo.u:7.5-9: inner
\end{urbiscript}

\end{urbiscriptapi}


%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
