\documentclass[openright,twoside,12pt]{report}
 \usepackage{urbi-specs}
% Index.  Must be before loading hyperref.
\usepackage{index}
\makeindex
% \usepackage{indexing}


% CONFIGURATION
\title{\us 2.0 specifications}
\author{Kernel Team}

% DOCUMENT
\begin{document}

\maketitle
\tableofcontents

\chapter*{Introduction}

\section*{Status of this document}

This document defines the specifications of the \us language version
2.0. It defines the expected behavior from the \us interpreter, the
standard library, and the \sdk. It can be used to check that some code
is valid, or browse \us or \Cxx \api for a desired feature. Reading
some parts with no particular reason can also provide you with
advanced knowledge or subtleties about some \us aspects.

This document is however not fit to be used as a start guide to learn
\us, since it is not structured in a progressive manner and too
detailed. You can think of this document as a dictionary: one does not
% FIXME: give directions to the tutorials
learn a foreign language by reading a dictionary. The \us Tutorial, or
the live \us tutorial built in the interpreter are good introduction
to \us.

This document doesn't either aim at giving advanced programming
techniques. Its only goal is to define how the language and its
libraries behave.

\section*{\us}

\us is an interpreted language designed for highly interactive and
parallel behavior programming. %FILLME

\section*{Structure of this document}

This document is divided in four chapters:

\begin{description}
  \newcommand{\xitem}[2]{\item[\autoref{#1} --- #2]~\\}
\xitem{sec:tools}{Tools specifications}%
  Presentation and usage of the different tools available with the
  \urbi framework related to \us, such as the \urbi server, the
  command line client, \umake, \ldots

\xitem{sec:lang}{\us language specifications}%
  Core constructs of the language and their behavior.

\xitem{sec:stdlib}{\us standard library specifications}%
  Listing of all classes and methods provided in the standard library.

\xitem{sec:sdk}{\urbi \sdk specifications}%
  The \urbi software development kit that enable to
  interact with \urbi from \Cxx.
\end{description}

\input{tools}
\FloatBarrier
\input{lang}
\FloatBarrier
\input{stdlib}
\FloatBarrier

\chapter{\urbi SDK specifications}
\label{sec:sdk}

% I just don't know
% what to do
% with myself
\clearpage
\phantomsection % sinon hyperliens vers chapitre précédent.
\addcontentsline{toc}{chapter}{\indexname}
\printindex

\end{document}

% printing, operators, waituntil

% LocalWords:  Urbi monomethods netcat cosinus timestamp subscopes subscope lst
% LocalWords:  getSlot setSlot updateSlot slotNames removeSlot uid arg ok ko ss
% LocalWords:  elt baz quux FIXME getslot lookup protos proto addProto asString
% LocalWords:  removeProto locateSlot locateslot lhs evalArgAt evalArgs ary
% LocalWords:  mytag timeOut myEvent Tuple redefinable CallMessage memoization
% LocalWords:  prepended
