%% Copyright (C) 2010, 2011, Gostai S.A.S.
%%
%% This software is provided "as is" without warranty of any kind,
%% either expressed or implied, including but not limited to the
%% implied warranties of fitness for a particular purpose.
%%
%% See the LICENSE file for more information.

\section{UVar}

This class is used internally by the UObject middleware (see
\autoref{part:uobject}) to represent a variable that can be hooked in
\Cxx. Each slot on which a \Cxx \lstinline|urbi::UVar| exists contains an
instance of this class.

Instances of \lstinline|UVar| are mostly transparent, they appear as
the value they contain. Thus, since the \lstinline|UVar| evaluates to the
contained value, you must use \lstinline|getSlot| to manipulate the
\lstinline|UVar| itself.

\subsection{Construction}
To instantiate a new \lstinline|UVar|, pass the owner object and the slot
name to the constructor.
\begin{urbiscript}
UVar.new(Global, "x")|
Global.x = 5;
[000000001] 5
x;
[000000002] 5
\end{urbiscript}

\subsection{Prototypes}
\begin{refObjects}
\item[Object]
\end{refObjects}

\subsection{Slots}

\begin{urbiscriptapi}

\item[copy](<targetObject>, <targetName>)%
  Copy the \lstinline|UVar| to the slot \var{targetName} of object
  \var{targetObject}. Since the \lstinline|UVar| is using properties, you
  must use this method to copy it to an other location.

\item[hookChanged]%
  A Boolean that states whether future UVar instantiations support the
  \lstinline|changed| event.  Code like:

\begin{urbiunchecked}
at (headTouch.val->changed? if headTouch.val)
  tts.say("ouch");
\end{urbiunchecked}
  \noindent
  works, but costs one \lstinline|at| per UVar (whose cost is light in
  recent version of \usdk).  Set \refSlot{hookChanged} to false
  \emph{before} instantiating new UVars to disable this feature; you may
  also set it to true to re-enable the feature for UVars that will be
  instantiating afterward.  See also the environment variable
  \env{URBI\_UVAR\_HOOK\_CHANGED}, \autoref{sec:tools:env}.

\item[notifyAccess](<onAccess>)%
  Similar to the \Cxx \lstinline|UNotifyAccess|, calls \var{onAccess} each
  time the \lstinline|UVar| is accessed (read).

\begin{urbiscript}
var Global.counter = 0|
UVar.new(Global, "access")|
var accessHandle = &access.notifyAccess(closure() {
  Global.access = Global.counter += 1
})|
assert
{
  access == 1;
  access == 2;
  access == 3;
};
&access.removeNotifyAccess(accessHandle)|;
assert
{
  access == 3;
  access == 3;
};
\end{urbiscript}


\item[notifyChange](<onChange>)%
  Similar to the \Cxx \lstinline|UNotifyChange| for a non-owned UVar (see
  \autoref{sec:uobject:uvar-notify}), register \var{onChange} and call it
  each time this \lstinline|UVar| is written to.  Return an identifier that
  can be passed to removeNotifyChange to unregister the callback.

\begin{urbiscript}
UVar.new(Global, "y")|
var handle = Global.&y.notifyChange(closure() {
  echo("The value is now " + Global.y)
})|
Global.y = 12;
[00000001] *** The value is now 12
[00000002] 12
Global.&y.removeNotifyChange(handle)|;
Global.y = 13;
[00000003] 13
\end{urbiscript}

\item[notifyChangeOwned](<onChangeOwned>)%
  Similar to the \Cxx \lstinline|UNotifyChange| for an owned UVar (see
  \autoref{sec:uobject:uvar-notify}), register \var{onChange} and call it
  each time this \lstinline|UVar| is written to.  Return an identifier that
  can be passed to removeNotifyChange to unregister the callback.

\item[removeNotifyAccess](<id>)%
  Disable the notification installed as \var{id} by \refSlot{notifyAccess}.

\item[removeNotifyChange](<id>)%
  Disable the notification installed as \var{id} by \refSlot{notifyChange}.

\item[removeNotifyChangeOwned](<id>)%
  Disable the notification installed as \var{id} by
  \refSlot{notifyChangeOwned}.

\item[owned]%
  True if the \lstinline|UVar| is in \dfn{owned mode}, that is if it
  contains both a sensor and a command value.

\end{urbiscriptapi}

%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-master: "../urbi-sdk"
%%% ispell-dictionary: "american"
%%% ispell-personal-dictionary: "../urbi.dict"
%%% fill-column: 76
%%% End:
